Hierarchy Builder: algebraic~hierarchies made easy in Coq with Elpi

### 2020-05-04 hb.tex
fix category/relatedversion
Renaming A -> R, switching to code that compiles as such.

\category{}%optional, e.g. invited paper

\relatedversion{System Description}%optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversion{A full version of the paper is available at \url{...}.}

@@

Lemma addrC : forall R : Ring.type, commutative (@add R).                       #\label{demo:theory:state:addrC}#
Proof. (* Proof by Hankel 1867, in Appendix B *) Qed.

Definition Z_Monoid_axioms : Monoid_of_Type Z :=                         #\label{demo:theory:z:monoid:axioms}#
  Monoid_of_Type.Build Z 0 Z.add Z.add_assoc Z.add_0_l Z.add_0_r.

@@

Definition Z_Ring_axioms : Ring_of_Monoid Z :=                        #\label{demo:theory:z:ring:axioms}#
  Ring_of_Monoid.Build Z 1 Z.opp Z.mul
    Z.add_opp_diag_l Z.add_opp_diag_r Z.mul_assoc Z.mul_1_l Z.mul_1_r
    Z.mul_add_distr_r Z.mul_add_distr_l.

@@

\section{Proof of addrC}
\label{sec:proof-addrc}
\begin{coqcode}
Lemma addrC {R : Ring.type} : commutative (@add R).
Proof.
have innerC (a b : R) : (a + b) + (a + b) = (a + a) + (b + b).
  by rewrite -[a+b]mul1r -mulrDl mulrDr !mulrDl !mul1r.
have addKl (a b c : R) : a + b = a + c -> b = c.
  apply: can_inj (add a) (add (-a)) _ _ _.
  by move=> x; rewrite addrA addNr add0r.
have addKr (a b c : R) : b + a = c + a -> b = c.
  apply: can_inj (add ^~ a) (add ^~ (-a)) _ _ _.
  by move=> x; rewrite /= -addrA addrN addr0.
move=> x y; apply: addKl (x) _ _ _; apply: addKr (y) _ _ _.
by rewrite -!addrA [in RHS]addrA innerC !addrA.
Qed.
\end{coqcode}

@@

HB.builders Context (A : Type) (f : Ring_of_Monoid A).

  Lemma addrC : commutative (add : A -> A -> A).
  Proof.
  have innerC (a b : A) : a + b + (a + b) = a + a + (b + b).
    by rewrite -[a+b]mul1r -mulrDl mulrDr !mulrDl !mul1r.
  have addKl (a b c : A) : a + b = a + c -> b = c.
    apply: can_inj (add a) (add (opp a)) _ _ _.
    by move=> x; rewrite addrA addNr add0r.
  have addKr (a b c : A) : b + a = c + a -> b = c.
    apply: can_inj (add ^~ a) (add ^~ (opp a)) _ _ _.
    by move=> x; rewrite /= -addrA addrN addr0.
  move=> x y; apply: addKl (x) _ _ _; apply: addKr (y) _ _ _.
  by rewrite -!addrA [in RHS]addrA innerC !addrA.
  Qed.

  (* Builder to the bottom right mixin. *)
  Definition to_AbelianGroup_of_Monoid :=
    AbelianGroup_of_Monoid.Build A opp addrC addNr.
  HB.instance A to_AbelianGroup_of_Monoid.

  (* Builder to the top right factory, which is compiled to the bottom left mixin. *)
  Definition to_Ring_of_AbelianGroup := Ring_of_AbelianGroup.Build A one mul
    mulrA mul1r mulr1 mulrDl mulrDr.
  HB.instance A to_Ring_of_AbelianGroup.

>>>>

\category{System Description}%optional, e.g. invited paper

%\relatedversion{System Description}%optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversion{A full version of the paper is available at \url{...}.}

@@

Lemma addrC {R : Ring.type} : commutative (@add R).                       #\label{demo:theory:state:addrC}#
Proof. (* Proof by Hankel 1867, in Appendix B *) Qed.

Definition Z_Monoid_axioms : Monoid_of_Type Z :=                         #\label{demo:theory:z:monoid:axioms}#
  Monoid_of_Type.Build Z 0%Z Z.add Z.add_assoc Z.add_0_l Z.add_0_r.

@@

Definition Z_Ring_axioms : Ring_of_Monoid Z :=                        #\label{demo:theory:z:ring:axioms}#
  Ring_of_Monoid.Build Z 1%Z Z.opp Z.mul
    Z.add_opp_diag_l Z.add_opp_diag_r Z.mul_assoc Z.mul_1_l Z.mul_1_r
    Z.mul_add_distr_r Z.mul_add_distr_l.

@@

\section{Proof of addrC}
\label{sec:proof-addrc}
\begin{coqcode}
Lemma addrC {R : Ring.type} : commutative (@add R).
Proof.
have innerC (a b : R) : a + b + (a + b) = a + a + (b + b).
  by rewrite -[a+b]mul1r -mulrDl mulrDr !mulrDl !mul1r.
have addKl (a b c : R) : a + b = a + c -> b = c.
  apply: can_inj (add a) (add (opp a)) _ _ _.
  by move=> x; rewrite addrA addNr add0r.
have addKr (a b c : R) : b + a = c + a -> b = c.
  apply: can_inj (add ^~ a) (add ^~ (opp a)) _ _ _.
  by move=> x; rewrite /= -addrA addrN addr0.
move=> x y; apply: addKl (x) _ _ _; apply: addKr (y) _ _ _.
by rewrite -!addrA [in RHS]addrA innerC !addrA.
Qed.
\end{coqcode}

@@

HB.builders Context (R : Type) (f : Ring_of_Monoid R).

  Lemma addrC : commutative (add : R -> R -> R).
  Proof. (* Exactly the same as in Appendix B. *)
  have innerC (a b : R) : a + b + (a + b) = a + a + (b + b).
    by rewrite -[a+b]mul1r -mulrDl mulrDr !mulrDl !mul1r.
  have addKl (a b c : R) : a + b = a + c -> b = c.
    apply: can_inj (add a) (add (opp a)) _ _ _.
    by move=> x; rewrite addrA addNr add0r.
  have addKr (a b c : R) : b + a = c + a -> b = c.
    apply: can_inj (add ^~ a) (add ^~ (opp a)) _ _ _.
    by move=> x; rewrite /= -addrA addrN addr0.
  move=> x y; apply: addKl (x) _ _ _; apply: addKr (y) _ _ _.
  by rewrite -!addrA [in RHS]addrA innerC !addrA.
  Qed.

  (* Builder to the bottom right mixin. *)
  Definition to_AbelianGroup_of_Monoid :=
    AbelianGroup_of_Monoid.Build R opp addrC addNr.
  HB.instance R to_AbelianGroup_of_Monoid.

  (* Builder to the top right factory, which is compiled to the bottom left mixin. *)
  Definition to_Ring_of_AbelianGroup := Ring_of_AbelianGroup.Build R one mul
    mulrA mul1r mulr1 mulrDl mulrDr.
  HB.instance R to_Ring_of_AbelianGroup.
